\chapter*{Abstract}

%
Software providers handling sensitive data must verify that their applications conform to a wide breadth of \policies.
%
For example, to comply with the GDPR, applications must delete a user's personal data on request.
% 

\sys{} is a static analyzer that helps developers verify the compliance of their Rust applications.
%
\Devs{} write custom \policies{} for their application, which \sys{} checks against the implementation.
%
We imagine that \devs{} use \sys{} as part of their normal code review processes, allowing them to catch bugs before deployment.
%
\sys{} must therefore be \emph{practical}: it should be fast, robust to source code changes, and straightforward to use.

Writing a \sys{} policy, however, is not straightforward.
%
Policies are Rust programs which reason about 
data flow between particular source code entities (e.g., arguments or types).
%
As such, policies are closely tied to the application's implementation,
which makes them difficult to write and inaccessible to all those but the application's developers.

\syslang{} is a new \policylang{} that seeks to reduce the developer effort required to apply \sys{} to applications.
%
It can specify a wide range of \policies{} in an intuitive, 
controlled natural language format.
%
\Devs{} compile their \syslang{} policies to \sys{} Rust policies,
which they can then run against their applications to verify compliance.
%

We evaluate \syslang{} on seven real-world Rust applications and find that it can
express a wide variety of \policies{}, 
it incurs low \dev{} effort,
and that it is executes quickly enough for use in a production setting.
